### 2019-09-18 main.tex

\documentclass[11pt, letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{eso-pic}
\usepackage{soul}
\usepackage{xcolor}

\usepackage{gfsdidot} % Use the GFS Didot font: http://www.tug.dk/FontCatalogue/gfsdidot/
\usepackage{microtype} % Improves typography

\usepackage{fancyhdr}
% \usepackage{pagecounting}
% \usepackage[dvips]{color}

\newcommand{\illinoislogo}{\includegraphics[height=0.5in]{images/illinoislogo.png}}
\newcommand{\anslogo}{\hspace{-4mm}\includegraphics[height=0.5in]{images/ansuiuc.png}}
\newcommand{\anssite}{ans.npre.illinois.edu}
% \usepackage{background}
\usepackage{float}
\usepackage[top=2cm, bottom=2cm, outer=2cm, inner=2cm]{geometry}
\setcounter{tocdepth}{3}
% \setcounter{secnumdepth}{3}
\usepackage{environ}
\NewEnviron{letter}{%
  \input{structure}
  \opening
  \closing
}

\graphicspath{{./images/}}

% Creates the flow chart for conference management
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{chair} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
\tikzstyle{director} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=orange!30]
\tikzstyle{coordinator} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=white!30]
\tikzstyle{arrow} = [thick,->,>=stealth]

\begin{document}

\begin{titlepage}
\AddToShipoutPictureBG*{\includegraphics[angle=90, width=\paperwidth,height=\paperheight]{earth1.png}};

   \begin{center}

       \vspace*{1cm}

       \Huge\color{white}\textbf{Saving the World One Atom at a Time}

       \vspace{0.5cm}
       Nuclear, Plasma, and Radiological Engineering

       \vspace{1.5cm}

       \Large\textbf{Presented by ANS at the University of Illinois Urbana-Champaign}

       \vfill
       \vspace{0.8cm}

       \includegraphics[width=0.4\textwidth]{ans_sc_logo3.png}
   \end{center}
\end{titlepage}

\clearpage
\newpage

\pagestyle{fancy}
\lhead{\anslogo}
\rhead{\illinoislogo}

[Letter from the Chairs goes here!]
\input{letter}

\newpage
\tableofcontents
\newpage

\input{theme}
\input{uiuc}
\input{conflog}
\input{proglog}
\input{management}
\input{budget}

\end{document}