### 2019-09-25 banking.tex
updates management

\subsection{Banking}

In order to properly handle all of the conference expenses, two accounts will be used. One account will be with the national ANS headquarters, and the second will be a local checking account through Busey Bank. Our student section currently holds a checking account with Busey Bank and has developed a working relationship with them. Similar to previous conferences, the national ANS account will be the primary account due to their previous experience in handling conference funds and their 501(c)(3) tax exemption status. Although an account with the national ANS organization is currently not maintained by the UIUC local chapter, should this bid be selected, the opening of an account would happen almost immediately. An established local checking account with Busey Bank allows for convenience when using them as a secondary account. The student section account has been maintained for a number of years, allowing for adequate knowledge of Busey’s policies and a stable relationship to be established with the local branch. A new account would be opened, such that the student chapter funds and the conference funds are completely separate and require different oversight. This account will be used for small expenses that can occur during the conference. Although the ANS-managed account could be used for such purposes, the presence of a local branch allows for more flexibility if a purchase becomes time sensitive. Should the additional Busey Bank account be unobtainable, an account would be established with Chase instead. If ANS National desires to manage all funds, accommodations will be made to consolidate the funds.

\subsubsection{Financial Oversight}
Financial integrity must be maintained when providing a conference of this size. To do so, diligent oversight will be practiced for all transactions related to the conference. Expense requests will be required for all transactions. These requests must carefully outline the reason for the purchase and the total cost. If the purchase is reoccuring, automation will be required at the time of request. These requests will require the approval of both conference chairs in addition to the financial director. Only the conference chairs and the financial director will have authority, assuming the previously mentioned permission, to complete transactions on the Busey Bank account. To ensure transparency, the financial director will update a public record containing all transactions.