### 2019-09-25 liability.tex

\section{Liability}
In the process of hosting a conference, especially a student conference, great care must be placed in assuring the protection of all involved parties, particularly ANS National and the hosting university,  should a concern for liability arise. The following sections outline such cases.

\subsection{Hotels}
The Illini Union and the Garden Hotel in Urbana (soon to be the Radisson) are not responsible for lost, stolen, or damaged items. Damages to their premise will be charged to those responsible. These terms are to be agreed upon contractually on a guest by guest basis as rooms become occupied. Hotels are covered through their own insurance policies.

\subsection{Transportation}

\subsubsection{Peoria Charter}
Peoria Charter is registered with the United Motorcoach Association, International Motorcoach Group, US Department of Transportation, the cities of Champaign-Urbana, and is fully insured in the State of Illinois. They also hold a certificate of registration from the University of Illinois board. Peoria Charter assumes responsibility for the safety of their passengers.

\subsubsection{UIUC Carpool}
While using UIUC rental shuttles, employees of the University with valid driver’s license will be drivers. The State of Illinois Self-Insured Motor Vehicle Liability Plan provides liability coverage to full-time, part-time, and student employees or student trustees operating a University motor vehicle in the course of their employment and in accordance with State of Illinois and University policies and procedures. The State of Illinois may accept liability to others harmed by a University employee's negligence, but occupants of University motor vehicles are not provided any medical payments. Any personal property belonging to drivers and/or occupants is the responsibility of those individuals. Neither the University nor the State of Illinois provides any insurance coverage for personal property.

\subsubsection{CU-MTD}
Though CU-MTD is the recommended mode of transportation while on campus, they are not affiliated with the conference. They are, however, liable for the safety of their passengers.

\subsection{Dining}
University Catering and the Garden Hotel in Urbana (soon to be the Radisson) assume all liability regarding foodborne illness derived from their respective events.

\subsection{Bank Account and Expenses}
ANS claims liability for unpaid expenses with primary banking orchestrated through ANS National. All receipts will be managed by the Financial Chair It is thus imperative to save expenditure receipts and review all contracts involving the exchange of money carefully as to avoid unexpected expenses.

\subsection{Release of Information}
Since personal information - such as credit card numbers, addresses, and contact information - is necessarily gathered during registration, it is imperative that any and all leaks be avoided. ANS claims liability for leaks of secure information. Such information provided by attendees to secondary services do not fall under the liability of the conference.

\subsection{Disabilities}
UIUC is proudly a fully ADA compliant university. As such, all events hosted on campus will be compliant with ADA regulations. There can be no guarantee regarding the compliance of off-campus establishments.

\subsection{Alcohol}
During conference hosted or sponsored events at which alcohol is being served, only those at or above the age of 21 years who can produce a valid ID will be served. Establishments licensed to serve alcohol are liable for ensuring the age of their customers. Catering services reserve the right to deny service to those who cannot produce a valid form of ID regarding the sale of alcohol. Special care will be taken at events serving alcohol to ensure safe consumption and avoid cases of illness due to over-consumption, and to ensure individuals deemed too drunk will no longer be served and can return to their rooms safely.
Patrons of Champaign-Urbana bars must be 19 years old to enter. It is the responsibility of those establishments to verify the age of their customers before serving alcohol. ANS does not assume liability for underage consumption of alcohol.