### 2019-09-27 conflict.tex
add conflict calendars

\section{Appendix: Conflict Calendar}

\begin{figure}[H]
	\centering
	\textbf{March 2021}\par\medskip
	\includegraphics[width=\textwidth, height=9cm]{march-conflict.png}
	\caption{Potential Date Conflicts in March 2021}
\end{figure}

\vspace{1cm}

\begin{figure}[H]
	\centering
	\textbf{April 2021}\par\medskip
	\includegraphics[width=\textwidth, height=9cm]{april-conflict.png}
	\caption{Potential Date Conflicts in April 2021}
\end{figure}