### 2019-09-28 main.tex
changes title

\documentclass[10pt, letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{lipsum}
\usepackage{wrapfig}
\usepackage{eso-pic}
\usepackage{soul}
\usepackage{xcolor}
\usepackage[parfill]{parskip}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{gfsdidot} % Use the GFS Didot font: http://www.tug.dk/FontCatalogue/gfsdidot/
\usepackage{microtype} % Improves typography
\usepackage[toc,page]{appendix} % creates an appendix
\usepackage{fancyhdr}
\usepackage{pdfpages}
\usepackage{multirow}
\usepackage{array}
% \usepackage{pagecounting}
% \usepackage[dvips]{color}
\newcommand{\STAB}[1]{\begin{tabular}{@{}c@{}}#1\end{tabular}}
\newcommand{\illinoislogo}{\includegraphics[height=0.5in]{images/block-I.png}}
\newcommand{\anslogo}{\hspace{-4mm}\includegraphics[height=0.5in]{images/ansuiuc.png}}
\newcommand{\anssite}{ans.npre.illinois.edu}
% \usepackage{background}
\usepackage{float}
\usepackage[top=2cm, bottom=2cm, outer=2cm, inner=2cm]{geometry}
\setcounter{tocdepth}{3}
% \setcounter{secnumdepth}{3}
\usepackage{environ}
\NewEnviron{letter}{%
  \input{structure}
  \opening
  \closing
}

\graphicspath{{./images/}}

% Creates the flow chart for conference management
\usepackage{varwidth}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{chair} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black,thick, fill=blue!50, execute at begin node={\begin{varwidth}{2.5cm}},execute at end node={\end{varwidth}}]
\tikzstyle{director} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=orange!70, execute at begin node={\begin{varwidth}{2.5cm}},execute at end node={\end{varwidth}}]
\tikzstyle{coordinator} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=white!30, execute at begin node={\begin{varwidth}{2.5cm}},execute at end node={\end{varwidth}}]
% \tikzstyle{arrow} = [thick,--,>=stealth]

\begin{document}
% ======================================================================================================
% DO NOT EDIT THIS SECTION
% ======================================================================================================
\begin{titlepage}
\AddToShipoutPictureBG*{\includegraphics[angle=90,width=\paperwidth,height=\paperheight]{earth1.png}}

   \begin{center}

       % \vspace{0.15cm}

       \Huge\color{white}\textbf{Saving the World One Atom at a Time}

       % \vspace{0.1cm}
       % Nuclear, Plasma, and Radiological Engineering
       % \includegraphics[width=0.4\textwidth]{ans_sc_logo3.png}

       \vfill
       \vspace{8.5cm}

       \includegraphics[width=0.4\textwidth]{ans_sc_logo3.png}

       \vspace{1cm}

       \Large\textbf{Presented by ANS at the University of Illinois Urbana-Champaign}

   \end{center}
\end{titlepage}
% ======================================================================================================
% ======================================================================================================
% ======================================================================================================

\clearpage
\newpage

\setlength{\headheight}{25pt}
\pagestyle{fancy}
\lhead{\anslogo}
\rhead{\illinoislogo}
% \renewcommand{\headrulewidth}{0pt} % no line in header area
% \fancyhead{}

\includepdf[pages=-]{./documents/chair_letter.pdf}

\newpage
\tableofcontents
\newpage

\input{theme}
\newpage
\input{uiuc}
\newpage
\input{conflog}
\input{proglog}
\newpage
\input{management}
\newpage
\input{website}
\newpage
\input{budget}
% \newpage
\input{liability}
\newpage
% Appendices
\appendix
\input{conflict}
\input{judging}
\input{schedule}
\input{building}
\input{locations}
\input{staffreq}
\input{facultyletter}
\input{petition}

\end{document}