### 2019-09-30 staffreq.tex
adds staff requirements

\section{Appendix: Staffing Requirements}
\label{appendix:staff}

\begin{table}[H]
	\caption{Staffing Requirements by Day}
	\label{table:staff}
	\centering
	\begin{tabular}{ccc}
	\hline\hline
	\textbf{Event} & \textbf{Description} & \textbf{Volunteers} \\
	\hline\hline
	& \textbf{Thursday} & \\
	\hline
	Registration & 3/shift, 5 2-hr shifts & 15\\
	Student HQ & 2/shift, 5 2-hr shifts & 10\\
	Professional's Lounge & 1/shift, 5 2-hr shifts & 5\\
	Workshops & 1/workshop for technical assistance & 4\\
	Bussed Tours & 1/tour for logistic assistance & 4\\
	Campus Tours & 4 tour groups, 2 per group & 8\\
	Dinner & 4 all purpose & 4\\
	Social & 2 all purpose & 2\\
	Career Fair Setup & 2 to assist recruiters & 2 \\
	\hline	
	& Total & 54\\
	\hline
	& \textbf{Friday} & \\
	\hline
	Registration & 2/shift, 5 2-hr shifts & 10\\
	Student HQ & 2/shift, 5 2-hr shifts & 10\\
	Professional's Lounge & 1/shift, 5 2-hr shifts & 5\\
	Technical Sessions & Session MCs & 24\\
	Panels & 1/panel for technical assistance & 4\\
	Dinner & 4 all purpose & 4\\
	Social & 2 all purpose & 2\\
	\hline
	& Total & 59\\
	\hline
	& \textbf{Saturday} & \\
	\hline
	Registration & 2/shift, 5 2-hr shifts & 10\\
	Student HQ & 2/shift, 5 2-hr shifts & 10\\
	Professional's Lounge & 1/shift, 5 2-hr shifts & 5\\
	Technical Sessions & Session MCs & 24\\
	Panels & 1/panel for technical assistance & 4\\
 	Poster Session & 2 all purpose & 2\\
	Dinner & 4 all purpose & 4\\
	Social & 2 all purpose & 2\\
	\hline
	 & Total & 61\\
	\hline
	\end{tabular}
\end{table}