### 2019-10-01 staffreq.tex
adds reporting relationship to staffreq

\section{Appendix: Staffing Requirements}
\label{appendix:staff}

\begin{table}[H]
	\caption{Staffing Requirements by Day}
	\label{table:staff}
	\centering
	\begin{tabular}{ccc|c}
	\hline\hline
	\textbf{Event} & \textbf{Description} & \textbf{Volunteers}& \textbf{Reports To} \\
	\hline\hline
	& \textbf{Thursday} & &\\
	\hline
	Registration & 3/shift, 5 2-hr shifts & 15& Registration Coordinator\\
	Student HQ & 2/shift, 5 2-hr shifts & 10& Program Coordinator\\
	Professional's Lounge & 1/shift, 5 2-hr shifts & 5& Program Coordinator\\
	Workshops & 1/workshop for technical assistance & 4& Workshops Coordinator\\
	Bussed Tours & 1/tour for logistic assistance & 4& Tours Coordinator\\
	Campus Tours & 4 tour groups, 2 per group & 8& Tours Coordinator\\
	Dinner & 4 all purpose & 4& Hospitality Coordinator\\
	Social & 2 all purpose & 2& Hospitality Coordinator\\
	Career Fair Setup & 2 to assist recruiters & 2 & Career Fair Coordinator\\
	\hline	
	& Total & 54&\\
	\hline
	& \textbf{Friday} & &\\
	\hline
	Registration & 2/shift, 5 2-hr shifts & 10& Registration Coordinator\\
	Student HQ & 2/shift, 5 2-hr shifts & 10& Program Coordinator\\
	Professional's Lounge & 1/shift, 5 2-hr shifts & 5& Program Coordinator\\
	Technical Sessions & Session MCs & 24& Sessions Coordinator\\
	Panels & 1/panel for technical assistance & 4& Workshops Coordinator\\
	Dinner & 4 all purpose & 4& Hospitality Coordinator\\
	Social & 2 all purpose & 2& Hospitality Coordinator\\
	\hline
	& Total & 59&\\
	\hline
	& \textbf{Saturday} & &\\
	\hline
	Registration & 2/shift, 5 2-hr shifts & 10& Registration Coordinator\\
	Student HQ & 2/shift, 5 2-hr shifts & 10& Program Coordinator\\
	Professional's Lounge & 1/shift, 5 2-hr shifts & 5& Program Coordinator\\
	Technical Sessions & Session MCs & 24& Sessions Coordinator\\
	Panels & 1/panel for technical assistance & 4& Workshops Coordinator\\
 	Poster Session & 2 all purpose & 2& Sessions Coordinator\\
	Dinner & 4 all purpose & 4& Hospitality Coordinator\\
	Social & 2 all purpose & 2& Hospitality Coordinator\\
	\hline
	 & Total & 61&\\
	\hline
	\end{tabular}
\end{table}