### 2019-10-01 conftrans.tex
makes various updates

\subsubsection{Getting to the Conference}
% The primary hotel, Garden Hotel Urbana, is approximately two miles from the Union and outside of reasonable walking distance. Thus we have budgeted to include transportation provided by Peoria Charter. We intend to have busses shuttling continuously between the Union and the Garden Hotel Urbana. For those who desire more flexibility and freedom, there is CU-MTD bus stop adjacent to the hotel that has a bus every 10 minutes. This bus route, the 22 Illini, can take attendees directly to the Union from the hotel. It is also one of the main campus bus routes and can take travelers almost anywhere on or around campus. Lyft and Uber are available as well for attendees that choose to use those services. A map of the bus route is in Appendix E: Campus Map of Locations on page \pageref{appendix:map}.
Each day of the conference, there will be four, 55 passenger, busses from Peoria Charter outside the hotel by 7:00 AM. They will leave as they fill up and be staggered such that there will be two busses arriving at either the Union or the hotel every ten minutes. These busses will run throughout the day from 7:00 AM to 5:00 PM. At 5:00 PM the busses will begin transporting attendees to the dinner venue (except on Thursday night when the dinner is located at the Garden Hotel) until 6:00 PM. Busses will be available at 8:00 PM until 9:00 PM to transport attendees either back to the hotel or to the evening social. The final rounds of busses will be optionally available to take attendees from the social to the hotel at 10pm. For those who desire more flexibility and freedom, there is CU-MTD bus stop adjacent to the hotel that has a bus every 10 minutes. This bus route, the 22 Illini, can take attendees directly to the Union from the hotel. It is also one of the main campus bus routes and can take travelers almost anywhere on or around campus. Lyft and Uber are available as well for attendees that choose to use those services. A map of the bus route is in Appendix E: Campus Map of Locations on page \pageref{appendix:map}.
\begin{figure}[H]
	\centering
	\label{figure:busses}
	\includegraphics[width=\textwidth]{transportation.png}
	\caption{Summary of Bus Transportation}
	\subcaption*{Same color refers to same set of busses.}
\end{figure}