### 2019-10-01 conflict.tex

\section{Appendix: Conflict Calendar}
\label{appendix:conflict}

\begin{figure}[H]
	\centering
	\textbf{March 2021}\par\medskip
	\includegraphics[width=\textwidth, height=9cm]{march-conflict.png}
	\caption{Potential Date Conflicts in March 2021}
\end{figure}

% \vspace{1cm}

\begin{figure}[H]
	\centering
	\textbf{April 2021}\par\medskip
	\includegraphics[width=\textwidth, height=9cm]{april-conflict.png}
	\caption{Potential Date Conflicts in April 2021}
	\subcaption*{Two primary (Date 1) conference dates are provided to avoid conflict with the annual UIUC Mom's Weekend}
\end{figure}