### 2020-11-19 addendum.tex
adds biographies and formatted images

\section*{Proposal Updates for 2020}

A few changes to the original ANS-UIUC proposal in response to the new conference date in 2022. Most of the changes involve updates to conference management structure, other changes are minimal but will be noted here.

\subsection*{Planning Committee Changes}

An overview of the changes to the planning committee are shown in Figure \ref{fig:newmgmt}. Notably, the roles of \textit{Technical Subcommittee Chair} and \textit{Logistics Subcommittee Chair} have been absorbed by the Technical Chair and the General Chair, respectively. The responsibilities of the  \textit{Sponsorship Coordinator} have been absorbed by the Financial Chair.

\begin{figure}[H]
\begin{center}
	\begin{tikzpicture}
		\node (gen chair) [chair] {General Co-Chair (Sam Dotson)};
		\node (tech chair) [chair, right of = gen chair, xshift = 5cm] {Technical Co-Chair (Jeremy Mettler)};
		\node (div chair) [chair, left of = gen chair, xshift = -5cm] {Diversity Co-Chair (Amanda Bachmann)};

		\node (prog coord) [coordinator, below of = gen chair, xshift=-1.75cm, yshift = -2cm] {Program Coordinator (Dilan K.)};
    \node (med coord) [coordinator, below of = div chair, xshift = -0.5cm, yshift = -2.0cm] {Media Coordinator (Nathan Ryan)};
		\node (hat coord) [coordinator, below of = prog coord, yshift = -2cm] {Hotels and Transport (Gavin Davis)};
		\node (hoa coord) [coordinator, below of = hat coord, yshift = -2cm] {Hospitality and Accomodations\\ (Nataly Panczyk)};
		\node (tours coord) [coordinator, below of = hoa coord, yshift = -2cm] {Tours Coordinator (Jake Mitstifer)};

		\node (work coord) [coordinator, below of = tech chair, xshift=0.5cm, yshift = -2cm] {Workshops Coordinator\\(Roberto Fairhurst)};
		\node (sesh coord) [coordinator, below of = work coord, yshift = -2cm] {Sessions Coordinator (Joshua Hoffman)};
		\node (cf coord) [coordinator, below of = sesh coord, yshift = -2cm] {Career Fair Coordinator (Greg Golba)};

		\node (fin chair) [director, below of = tech chair, xshift = -3.75cm, yshift = -2cm] {Financial Chair\\(Stephen Armstrong)};
		\node (reg coord) [coordinator, below of = fin chair, yshift = -2cm] {Registration Coordinator\\(Jasmine Dinari)};

		\draw  (gen chair) -- (tech chair);
		\draw  (gen chair) -- (div chair);

		\draw  (gen chair) -- (prog coord);
		\draw  (prog coord) -- (hat coord);
		\draw  (hat coord) -- (hoa coord);
		\draw  (hoa coord) -- (tours coord);
		\draw  (gen chair) -- (med coord);
		\draw  (tech chair) -- (fin chair);
		\draw  (gen chair) -- (fin chair);
		\draw  (fin chair) -- (reg coord);
		\draw  (tech chair) -- (work coord);
		\draw  (work coord) -- (sesh coord);
		\draw  (sesh coord) -- (cf coord);

	\end{tikzpicture}
  \caption{Note: The hierarchy of responsibility is denoted by color, not necessarily by the position on the chart.}
  \label{fig:newmgmt}
\end{center}
\end{figure}

% ===========================================================================
\subsection{Updated Biographies}
\input{new_biographies}