### 2020-11-19 management.tex

\section{Conference Management}

The Conference Planning Committee is separated into three categories: Co-Chairs, Directors, Coordinators. The three co-chairs are responsible for setting up major milestones and ensuring that those milestones are met. Together they oversee all three subcommittees. They will serve as the primary contacts between subcommittee chairs and the faculty as well as professionals. The co-chairs have the final word on all conference decisions. They serve as the faces of the conference. In this proposal we have decided to add a novel Co-Chair position: Diversity Co-Chair. This addition reflects our goal of celebrating the people behind the science and make this conference experience accessible to as many people as possible.
The Directors are in charge of specialized subcommittees and are responsible for making decisions in those areas. They must communicate with their subcommittees and with other directors to make sure there are no conflicting decisions. The coordinators work in an even more specialized area and help the directors by handling a single task.\\

% \vspace{1cm}

% ===============================================
% This is the hierarchy breakdown
% ===============================================
% \begin{center}
% 	\begin{tikzpicture}[
% 		chair/.style={rectangle, text centered, draw,fill=blue!20, execute at begin node={\begin{varwidth}{9em}},
%    		execute at end node={\end{varwidth}}},
%    		committee/.style={rectangle,draw,fill=green!40, text centered, execute at begin node={\begin{varwidth}{7em}},
%    		execute at end node={\end{varwidth}}},
%    		coordinator/.style={rectangle,draw,fill=red!30, text centered, execute at begin node={\begin{varwidth}{7em}},
%    		execute at end node={\end{varwidth}}},
% 		grandchild/.style={grow=down,xshift=1em,anchor=west,
% 		edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)}, execute at begin node={\begin{varwidth}{8em}},
%    		execute at end node={\end{varwidth}}},
% 		first/.style={level distance=12ex},
% 		second/.style={level distance=24ex},
% 		third/.style={level distance=36ex},
% 		level 1/.style={sibling distance=11em}]
% 	    % Parents
% 	    \coordinate
% 	      child[grow=left] {node[chair,anchor=east]{General Co-Chair \\ (Sam Dotson)}}
% 	      child[grow=right] {node[chair,anchor=west]{Technical Co-Chair\\ (Jeremy Mettler)}}
% 	      child[grow=down,level distance=3ex]
% 	    [edge from parent fork down]
% 	    % Children and grandchildren
% 	    child{node[committee] {Non-Technical Subcommittee}
% 	      child[grandchild,first] {node[chair]{Logistics Chair}}
% 	      child[grandchild,second] {node[coordinator]{Hotels and Transportation}}
% 	      child[grandchild,third] {node[coordinator] {Hospitality and Catering}}}
% 	    child {node[committee] {Media Subcommittee}
% 	      child[grandchild, first] {node[chair] {Publications Chair}}
% 	      child[grandchild, second] {node[coordinator]{Media}}
% 	      child[grandchild, third] {node[coordinator] {Program}}}
% 	    child{node[committee] {Financial Subcommittee}
% 	      child[grandchild,first] {node[chair]{Financial Chair}}
% 	      child[grandchild,second] {node[coordinator]{Registration}}
% 	      child[grandchild,third] {node[coordinator]{Sponsorship}}}
% 	    child {node[committee]{Technical Subcommittee}
% 	      child[grandchild,first] {node[chair]{Sessions Chair}}
% 	      child[grandchild,second] {node[coordinator]{Workshops}}
% 	      child[grandchild,third] {node[coordinator]{Career Fair}}};
% 	\end{tikzpicture}
% \end{center}
\begin{center}
	\begin{tikzpicture}
		\node (gen chair) [chair] {General Co-Chair (Sam Dotson)};
		\node (tech chair) [chair, right of = gen chair, xshift = 5cm] {Technical Co-Chair (Jeremy Mettler)};
		\node (div chair) [chair, left of = gen chair, xshift = -5cm] {Diversity Co-Chair (Nate Reid)};

		\node (log chair) [director, below of = gen chair, xshift = -3.75cm, yshift = -2cm] {Logistics Chair\\ (Natalie Gaughan)};
		\node (med coord) [coordinator, below of = gen chair, xshift = 0.0cm, yshift = -3cm] {Media Coordinator (Nathan Ryan)};
		\node (prog coord) [coordinator, below of = log chair, yshift = -2cm] {Program Coordinator (Dilan K.)};
		\node (hat coord) [coordinator, below of = prog coord, yshift = -2cm] {Hotels and Transport (Gavin Davis)};
		\node (hoa coord) [coordinator, below of = hat coord, yshift = -2cm] {Hospitality and Accomodations\\ (Brady Moran)};
		\node (tours coord) [coordinator, below of = hoa coord, yshift = -2cm] {Tours Coordinator (Anna Balla)};

		\node (tsc chair) [director, below of = tech chair, xshift = 2cm, yshift = -2cm] {Technical Sub-Committee Chair (Gwendolyn Chee)};
		\node (work coord) [coordinator, below of = tsc chair, yshift = -2cm] {Workshops Coordinator\\(Roberto Fairhurst)};
		\node (sesh coord) [coordinator, below of = work coord, yshift = -2cm] {Sessions Coordinator (Zach Jeckell)};
		\node (cf coord) [coordinator, below of = sesh coord, yshift = -2cm] {Career Fair Coordinator (Greg Golba)};

		\node (fin chair) [director, below of = tech chair, xshift = -2.25cm, yshift = -2cm] {Financial Chair\\(Jack Stahl)};
		\node (spon coord) [coordinator, below of = fin chair, yshift = -2cm] {Sponsorships Coordinator (Jimmy Shehee)};
		\node (reg coord) [coordinator, below of = spon coord, yshift = -2cm] {Registration Coordinator\\(Jasmine Dinari)};

		\draw  (gen chair) -- (tech chair);
		\draw  (gen chair) -- (div chair);
		\draw  (gen chair) -- (log chair);
		\draw  (log chair) -- (prog coord);
		\draw  (prog coord) -- (hat coord);
		\draw  (hat coord) -- (hoa coord);
		\draw  (hoa coord) -- (tours coord);
		\draw  (gen chair) -- (med coord);
		\draw  (tech chair) -- (fin chair);
		\draw  (gen chair) -- (fin chair);
		\draw  (tech chair) -- (tsc chair);
		\draw  (fin chair) -- (spon coord);
		\draw  (spon coord) -- (reg coord);
		\draw  (tsc chair) -- (work coord);
		\draw  (work coord) -- (sesh coord);
		\draw  (sesh coord) -- (cf coord);
		\draw  (div chair) -- (log chair);

	\end{tikzpicture}
\end{center}

% ===============================================
% ===============================================
% ===============================================

\newpage
\subsection{Position Responsibilities}

\begin{itemize}
	\item $\large\textbf{\textit{General Co-Chair}}$\\
	In addition to the responsibilities outlined above, the General Co-Chair oversees the
	Non-technical Subcommittee, the Media Coordinator, and the Financial Subcommittee. He works with the Technical Co-Chair in talking to sponsors, companies, and planning full committee meetings.
	\item $\large\textbf{\textit{Technical Co-Chair}}$\\
	In addition to the responsibilities outlined above, the Technical Co-Chair primarily oversees the
	Technical Subcommittee. Works with the General Co-Chair to plan full committee meetings and completes any remaining tasks to ensure that milestones are met.
	\item $\large\textbf{\textit{Diversity Co-Chair}}$\\
	In addition to the responsibilities outlined above, the Diversity
                Co-Chair is responsible for making sure the conference is as
                inclusive, diverse, and accessible as possible. This includes:
                contacting minority serving institutions to encourage their
                students to attend the conference, ensuring socials and dinners
                are accessible, verifying that the final list of speakers has
                adequate representation, engaging with the national Diversity and
                Inclusion in ANS Committee, and more. We believe the example set
                by UIUC will make this role a staple of future conferences.

	\item $\textbf{Non-Technical Subcommittee}$\\
	The Non-Technical Subcommittee oversees, arranges, and executes all actions related to hospitality, transportation, special events, and non-technical workshops and panels. This committee keeps the theme of the conference in mind when organizing all events. They ensure that the conference runs smoothly.
	\begin{itemize}
		\item[$\circ$] $\textbf{Logistics Chair}$\\
		The Logistics Chair is in charge of planning all tours and non-technical workshops and panels. Organizes speakers during dinners. Recruits student volunteers to staff events and sessions during the conference.
		Also responsible for keeping track of the subcommittee and reporting to the Co-Chairs.
		\item[$\circ$] $\textbf{Hotels and Transportation Coordinator}$\\
		Coordinates with ANS National to negotiate room rates and room blocks for hotels. Reserves buses for the
		necessary times and events. Works with the Hospitality and Catering Coordinator and the Logistics Chair.
		\item[$\circ$] $\textbf{Hospitality and Catering Coordinator}$\\
		Responsible for planning and organizing all catered meals for
                        the conference. That includes contacting the catering services, reserving the venues where meals are held, and making sure the venues are staffed.
		\item[$\circ$]$\textbf{Tours Coordinator}$\\
		The tours coordinator organizes tours and works with the transportation coordinator to ensure that transportation is available during the conference for these tours.
		\item[$\circ$]$\textbf{Program Coordinator}$\\
		The program coordinator creates programs for the conference, sets the conference schedule to minimize overlap between events and increase student involvement. Communicates with other coordinators and subcommittees about the schedule of events. Designs and purchases T-shirts for attendees. Arranges gift bags for attendees. Works with sponsorship coordinator to create gift bags.
	\end{itemize}
	\item $\textbf{Financial Subcommittee}$\\
	The Financial Subcommittee oversees, arranges, and executes all actions related to banking, sponsorship, registration, reimbursement, budgeting, and monetary exchanges. This committee works closely with the General and Technical Co-Chairs.
	\begin{itemize}
		\item[$\circ$] $\textbf{Financial Chair}$\\
		Manages the ANS Planning Committee account with Busey Bank and ANS National. The account coordinator is also responsible for keeping track of receipts, setting a budget for the committee, keeps track of all transactions,
		\item[$\circ$] $\textbf{Registration Coordinator }$
		Handles the registration for professional and student attendees. Communicates the number of attendees to the Program Coordinator.
		\item[$\circ$] $\textbf{Sponsorship Coordinator}$\\
		Assists the Financial Co-Chair with matters involving sponsorship as well as working closely with the Registration and Account Coordinators and the General and Technical Co-Chairs. Works with the Program Coordinator on gift bag items and works with the Career Fair Coordinator.
	\end{itemize}
	\item $\textbf{Technical Subcommittee}$\\
	The Technical Subcommittee works with the Technical Co-Chair to process student abstracts and set up
	technical workshops, panels and sessions.
	\begin{itemize}
		\item[$\circ$] $\textbf{Technical Subcommittee Chair}$\\
		 Also responsible for ensuring that judges understand the judging criteria, and organizes the award ceremony.
		Keeps track of subcommittee progress and reports to the General and Technical Co-chairs.
		\item[$\circ$] $\textbf{Sessions Coordinator}$\\
		Responsible for organizing presentation, poster sessions, and technical
		panels. Makes sure that all rooms are properly set up with equipment to ensure smooth technical sessions.
		\item[$\circ$] $\textbf{Workshops Coordinator}$\\
		Organizes workshops locations, times, staffing, costs, supplies, student enrollment
		and any other tasks required to have successful workshops. Also assists the Sessions Chair when needed.
		\item[$\circ$] $\textbf{Career Fair Coordinator}$\\
		Oversees staffing and support for the career fair as well as working with the Sponsorship
		Coordinator to ensure a successful career fair. Also assists the Sessions Chair when needed.
	\end{itemize}
	\item $\textbf{Media Coordinator}$\\
		The Media Coordinator is responsible for designing the conference website, constantly updating social media presence,
		and obtains information from other members of the planning committee for the website.
\end{itemize}

\newpage
\subsection{Planning Committee Biographies}
\input{biographies}

\subsection{Conflict Resolution}

Committee members should hold professionalism at the forefront of their composure in order to avoid and resolve issues amicably without the involvement of higher powers. As such, members are encouraged to settle conflicts without invoking this protocol. If an issue arises that immediately presents itself as overwhelming, the scope of the issue exceeds an individual’s ability to handle it, or the issue imposes certain implications that jeopardizes the mission of the planning committee, members should not hesitate to refer to this section. In the event of a conflict between members of the planning committee, a document for decision-making and
conflict resolution has been drafted and approved by the general committee. All General Committee members
are expected to abide by the resolution. Key points of the resolution are as follows:
\begin{enumerate}
	\item Subcommittees are encouraged to resolve conflicts internally and as democratically as possible. If an independent resolution cannot be reached, the Co-Chairs should be involved. The Co-Chairs have the final word on all decisions. If the Co-Chairs are unable to agree on a solution, the faculty advisor will be involved.
	\item Conflicts between individual committee members are to be resolved outside of the committee. Should such a conflict jeopradize the mission of the conference the Co-Chairs will be involved.
	\item Any cases of misconduct or negligence will be handled appropriately by the Co-Chairs.
	\item For extreme cases of misconduct or negligence, separate steps for the removal and replacement of a
	member are outlined for general members, Subcommittee Chairs, and Co-Chairs. These include
	a discussion with the offending member, consultation of the Faculty Advisor, and a hearing with the
	General Committee to decide if removal is necessary.
\end{enumerate}

\subsection{Staffing Requirements}
Staffing for room breakdowns and setups, registration, socials, workshops,
tours, etc.  will be supplied by either UIUC ANS members or students of the
NPRE department. While it is likely that members of our student chapter will
voluntarily fill all the staffing requirements of conference hosted events. If
not, we will rely on other student organizations for support including, but not
limited to: Arms Control and Domestic and International Security (ACDIS) and
Women in Engineering (WIE). Participating in conference events on a staff level
is a beneficial experience for any undergraduate who wants to become more
familiar with ANS or support their ANS Chapter. Throughout the conference,
interacting with registering students, driving groups to tour a scientific
facility, and turning rooms for technical sessions provide plentiful
opportunities for volunteers to interact with myriad members on a number of
levels. Thus, it will be an overall positive experience for the volunteers.
Quantitative needs are outlined in Appendix \ref{appendix:staff} on page
\pageref{appendix:staff}.

% ===============================================
% Milestones
% ===============================================
\newpage
\subsection{Milestones}

\begin{center}
\begin{longtable}{c |  c  c}
\hline\hline
\textbf{Deadline} & \textbf{Task} & \textbf{Responsible Committee}\\
\hline\hline

&\textbf{November}&\\
\hline\hline
11/21/19 & Announcement of 2021 Conference & \\
11/21/19 & Update milestones with lessons learned &\\
11/21/19 & Notify department of selection & Co-Chairs\\
11/22/19 & Confirm Conference Committee & Co-Chairs\\
11/29/19 & Finalize Conference Date & Co-Chairs\\
\hline\hline
&\textbf{December}&\\
\hline\hline
12/1/19 & Contact ANS National for support with hotel negotiations& Co-Chairs and Non-Technical\\
12/1/19 & Set up banking through Busey Bank and ANS National& Financial\\
12/3/19 & Confirm Conference Reservations & Non-Technical\\
12/3/19 & Finalize Facilities Reservations& Non-Technical\\
12/20/19 & Finalize Logo Design& Media\\
12/20/19 & Begin Designing Website and Social Media& Media\\
\hline\hline
&$\textbf{2020}$&\\
\hline\hline
&\textbf{January}&\\
\hline\hline
1/13/20 & Finalize meeting schedule& Co-Chairs\\
\hline\hline
&\textbf{February}&\\
\hline\hline
2/3/20 & Launch social media plan& Media\\
2/17/20 & Finalize list of potential speakers& Non-Technical\\
2/17/20 & Finalize technical topics& Technical\\
\hline\hline
&\textbf{March}&\\
\hline\hline
3/9/20 & Contact Speakers and Presenters& Non-Technical and Technical\\
3/26-28/20 & Attend Student Conference at NC State& All\\
3/28/20 & Request lessons learned from NC State& Co-Chairs\\
\hline\hline
&\textbf{April}&\\
\hline\hline
4/6/20 & Create sponsor letters and contact & General Co-Chair\\
4/13/20 & Prepare tour opportunities& Non-Technical\\
\hline\hline
&\textbf{May}&\\
\hline\hline
5/4/20 & Confirm tours and costs& Non-Technical\\
5/18/20 & Submit Progress Report to SSC& Co-Chairs\\
5/18/20 & Create a Call for Papers & Technical and Media\\
\hline\hline
&\textbf{June}&\\
\hline\hline
6/7-11/20 & Send Delegates to Annual Meeting in Arizona& Co-Chairs, All\\
\hline\hline
&\textbf{July}&\\
\hline\hline
7/6/20 & Confirm with all speakers and presenters& Technical\\
7/20/20 & Report on sponsorship& Financial\\
7/27/20 & Reassess budget& Financial\\
\hline\hline
&\textbf{August}&\\
\hline\hline
8/10/20 & Update with sponsors& Financial\\
8/24/20 & Submit second progress to SSC& Co-Chairs\\
\hline\hline
&\textbf{September}&\\
\hline\hline
9/14/20 & Updates on conference deadlines& Financial\\
9/21/20 & Finalize gift bag order& Non-Technical\\
\hline\hline
&\textbf{October}&\\
\hline\hline
10/12/20 & Confirm hotel blocks and contracts& Non-Technical\\
10/26/20 & Registration Opens& Financial\\
10/26/20 & Create and test online paper submission& Technical\\
\hline\hline
&\textbf{November}&\\
\hline\hline
11/15-19/20 & Attend Winter Conference 2020 in Chicago& Co-Chairs, All\\
11/24/20 & Finalize marketing material& Media\\
11/24/20 & Report from winter conference & Co-Chairs\\
\hline\hline
&\textbf{December}&\\
\hline\hline
12/4/20 & Reassess Budget & Financial \\
12/4/20 & Update website & Media \\
12/18/20 & Third progress report to SSC& Co-Chairs\\
12/18/20 & Send out call for papers& Media\\
\hline\hline
&$\textbf{2021}$&\\
\hline\hline
&\textbf{January}&\\
\hline\hline
1/4/20 & Confirm all judges, panelists, and speakers& Technical\\
1/18/20 & First paper deadline -- send to reviewers & Technical\\
1/22/20 & Recruit student volunteers& All\\
1/29/21 & Finalize tours& Non-Technical\\
1/29/21& Finalize conference transportation& Non-Technical\\
\hline\hline
&\textbf{February}&\\
\hline\hline
2/1/21 & Finalize program& Media\\
2/1/21 & Order gift bag items & Financial and Non-Technical\\
2/5/21 & Final Paper Deadline& Technical\\
2/9/21 & Extended Paper Deadline& \\
2/15/21 & Finalize Website& Media\\
2/19/21 & Finalize budget& Financial\\
\hline\hline
&\textbf{March}&\\
\hline\hline
3/22/21 & Print Programs& Media\\
3/26/21 & Finalize Awards& Technical\\
3/26/21 & Finalize Staff Schedule& Non-Technical\\
3/29/21 & Final progress report to SSC& Co-Chairs\\
\hline\hline
&\textbf{April}&\\
\hline\hline
4/1/21 & Prepare gift bags, print tags, and banners& Non-Technical\\
4/8-11/21 & Host ANS Student Conference 2021 & \\
4/22/21 & Return seed money& Financial\\
4/29/21 & Process Student Reimbursements& Financial\\
4/29/21 & Finalize financial report& Financial\\
4/29/21 & Submit conference report& Co-Chairs\\
\hline\hline
\end{longtable}
\end{center}